Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2014-12-22 parts/ccl.tex
continue transformation (encodings + examples + conclusion)

We introduced a new semantics to include priorities into the Process Hitting framework, which prove especially useful to model cooperations.
Then, we developed a method to efficiently perform a reachability analysis of a sequence of objectives in a restricted class of Process Hitting models,
but it is also useful to establish the reachability of a partial state.
This analysis is based on an under-approximation of the true reachability solutions.
%; however, the most usual cases can be handled.

We showed that the class of Process Hitting models that can be handled by the aforementioned method are equivalent to Asynchronous Discrete Networks, and therefore to Asynchronous Boolean Networks.
This allows to efficiently compute reachability results on large biological models provided that they are equivalent to Asynchronous Discrete Networks and that a translation from the original framework into a Process Hitting model is possible.
Such a translation for interaction graphs of Thomas modelling was proposed in~\cite{PMR10-TCSB}. % \towrite{[À garder ?] and is made possible by the Pint software}.

We also demonstrated that general Process Hitting models with arbitrary priority
classes can also be encoded in the particular class of two-prioritised Process Hitting.
Therefore, our static analysis is applicable to a broad range of automata
networks with priorities.

>>>>

In this paper, we focused on the Asynchronous Automata Networks (AANs),
which is a restriction of classical Automata Networks
and is equivalent to the Process Hitting with classes of priorities~\cite{FPMR13-CS2Bio}.
This formalism proves useful to model accurate Boolean gates, which was not possible
with the standard form of the Process Hitting,
and avoid an unwanted over-approximation of the dynamics.
Furthermore, we proposed an extension of this formalism with classes of priorities,
which prove convenient to abstract time parameters into these kinds of models,
or more simply to add preemption relations between actions.

Then, we developed a method to perform a reachability analysis
of a sequence of objectives in an AAN,
based on an under-approximation of the true reachability solutions.
The checked objectives can take the form of local states, global states
or partial states, making the analysis more convenient.
This method can be considered efficient,
because it is exponential in the number of possible solutions,
and each step is only polynomial in the size of the model.
Furthermore, AANs with classes of priorities can also be studied in this way,
at the cost of a translation that we gave in this paper.

Finally, we also showed that AANs are also equivalent to Asynchronous Discrete Networks,
that is, the multivalued version of the classical Asynchronous Boolean Networks
with discrete expression levels and evolution functions.
This framework is also equivalent to Thomas' models with Snoussi parameters
instead of evolution functions.
This especially allows to efficiently compute reachability results
on large biological models,
provided that they are equivalent to Asynchronous Discrete Networks,
which ensures that a translation to AANs is possible.
For example,
such a translation for generalized Interaction Graphs,
that is, Discrete Networks without evolution functions or parameters,
was proposed in~\cite{PMR10-TCSB}.
